---
title: "Statistical Analysis"
author: "Liow Yi Jia"
date: "3/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("AICcmodavg")
install.packages("xlsx")

library(broom)
library(MASS)
library("xlsx")
```

1) Normality Check
```{r}
outcome <- c("energy", "adj_protein", "adj_fat", "adj_carbohydrate", "fiber")

for(i in outcome){    
  shapiro <- shapiro.test(FOOD[[i]])
  print(i)
  print(shapiro)
}
```

2) Simple Linear Regression 
```{r}
FOOD <- na.omit(FOOD) # remove NAs from all variables

predictor <- c("region", "shop_convenience", "shop_frequency", "transportation", 
                     "time", "foodstuff", "bento", "homegrown")
covariate <- c("age", "sex", "alone", "household", "income", "education")
outcome <- c("energy", "adj_protein", "adj_fat", "adj_carbohydrate", "fiber")

variables <- c(predictor, covariate)

for (i in variables){ 
  model <- lm(formula(paste0("outcome ~", i[[1]])), data = FOOD) 
  print(summary(model)) 
} 

for (out in outcome){
  for (var in variables){
    model <- lm(formula(paste0(out, " ~ ", var)), data = FOOD) 
    print(out)
    print(summary(model))
  }
}

#lm_results <- lm(adj_protein ~ age, data = FOOD)
#summary(lm_results)
```

3) Multiple Linear Regression 
Question: Does food environment affect nutrient intake in elderly Japanese population?
```{r}
protein <- lm(adj_protein ~ shop_convenience + shop_frequency + transportation +
                time + foodstuff + bento + homegrown + age + sex + income +
                education, data = FOOD)
x <- tidy(protein)
augment(protein)
glance(protein)
summary(protein)

# multicollinearity check
vif_values <- vif(protein)
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue")

# for loop for MLR
for(i in outcome){
  mlr_result <- lm(FOOD[[i]] ~ shop_convenience + shop_frequency + transportation +
                time + foodstuff + bento + homegrown + age + sex + income +
                education, data = FOOD) 
  summary(mlr_result) # can store result in list
  print(i)
  print(summary(mlr_result))
}

# making a function to do MLR 
nutrition <- function(outcome){
  out_var <- formula(paste0(outcome, "~ region + shop_convenience + shop_frequency + 
                transportation + time + foodstuff + bento + homegrown + 
                age + sex + income + education"))
  mlr_result <- lm(out_var, data = FOOD)
  return(summary(mlr_result))
}

# example
#test2 <- function(outcome){
#  f1 <- formula(paste0(outcome, "~ shop_convenience"))
#  return(f1)
#}

energy_result <- nutrition("energy") %>% tidy()
protein_result <- nutrition("adj_protein") %>% tidy()
fat_result <- nutrition("adj_fat") %>% tidy()
carbohydrate_result <- nutrition("adj_carbohydrate") %>% tidy()
fiber_result <- nutrition("fiber") %>% tidy()
```

Export MLR results to Excel
```{r}
# Write the first data set in a new workbook
write.xlsx(energy_result, file = "~/Desktop/Food Environment.xlsx",
      sheetName = "Energy", append = FALSE)
# Add a second data set in a new worksheet
write.xlsx(protein_result, file = "~/Desktop/Food Environment.xlsx", 
           sheetName="Protein", append=TRUE)
# Add a third data set in a new worksheet
write.xlsx(fat_result, file = "~/Desktop/Food Environment.xlsx", 
           sheetName="Fat", append=TRUE)
# Add a fourth data set in a new worksheet
write.xlsx(carbohydrate_result, file = "~/Desktop/Food Environment.xlsx", 
           sheetName="Carbyhydrate", append=TRUE)
# Add a fifth data set in a new worksheet
write.xlsx(fiber_result, file = "~/Desktop/Food Environment.xlsx", 
           sheetName="Fiber", append=TRUE)
```


4) Variable Selection 
4-1) Stepwise Regression 
 - Combination of forward and backward selection of predictors
 - Add/remove predictors until preset significance level is met for all predictors
 
4-2) AIC (Akaike Information Criterion)
 - Find balance between the number of explanatory variables and goodness of fit
 - Metric to compare the fit between several regression models
 - Fit several regression models, compare AIC values of each model (the lower the better)
 
```{r}
# 4-1) Stepwise Regression
# select variables
food_4 <- food_env %>% select(region, sex, age, income, education, 
                             shop_convenience, shop_frequency, transportation, time, 
                             foodstuff, bento, homegrown, adj_protein)

str(food_4)

# make appropriate variable class
x <- c("region", "income", "education", "shop_convenience", "shop_frequency", "transportation",
       "foodstuff", "bento", "homegrown")
food_4[x] <- lapply(food_4[x], as.factor)

y <- c("age", "time")
food_4[y] <- lapply(food_4[y], as.integer)

food_4 <- na.omit(food_4) # remove NAs from all variables

# Fit the full model 
full_model <- lm(adj_protein ~., data = food_4)
# Stepwise regression model
step_model <- stepAIC(full_model, direction = "both", 
                      trace = FALSE)
summary(step_model)

################################################################################
# 4-2) AIC
# make different models
model1 <- lm(adj_protein ~ shop_convenience + transportation +
                time + foodstuff + bento + homegrown + age + sex + income +
                education, data = FOOD)
model2 <- lm(adj_protein ~ shop_convenience + shop_frequency + 
                time + foodstuff + bento + homegrown + age + sex + income +
                education, data = FOOD)
model3 <- lm(adj_protein ~ shop_convenience + shop_frequency + transportation +
                foodstuff + bento + homegrown + age + sex + income +
                education, data = FOOD)

#define list of models
models <- list(model1, model2, model3)

#specify model names
mod.names <- c('Model 1', 'Model 2', 'Model 3')

#calculate AIC of each model
aictab(cand.set = models, modnames = mod.names) #doesn't work - package doesn't load
```

